---
title: "Assignment 2"
author: Teja Tarapatla
output: html_document
date: "2024-02-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question:

Universal bank is a young bank growing rapidly in terms of overall customer acquisition.
The majority of these customers are liability customers (depositors) with varying sizes of
relationship with the bank. The customer base of asset customers (borrowers) is quite
small, and the bank is interested in expanding this base rapidly in more loan business. In
particular, it wants to explore ways of converting its liability customers to personal loan
customers.
A campaign that the bank ran last year for liability customers showed a healthy conversion
rate of over 9% success. This has encouraged the retail marketing department to devise
smarter campaigns with better target marketing. The goal is to use k-NN to predict whether
a new customer will accept a loan offer. This will serve as the basis for the design of a new
campaign.
The file UniversalBank.csv contains data on 5000 customers. The data include customer
demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan
campaign (Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the
personal loan that was offered to them in the earlier campaign.
Partition the data into training (60%) and validation (40%) sets.

```{r}
library(class)

```
```{r}
library(caret)
```

```{r}
univ_data <- read.csv("C:/Users/TEJA/Downloads/UniversalBank.csv")
dim(univ_data)
```

```{r}
head(univ_data)
```
```{r}
tail(univ_data)
```
```{r}
t(t(names(univ_data)))
```
```{r}
getwd()
```
```{r}
n_data <- univ_data[,-c(1,5)]
dim(n_data)
```
##60% should be used for training and 40% for validation. There are several ways to achieve this. Two methods will be taken into consideration. Before we split, let's first turn the category variables into dummy variables.

```{r}
n_data$Education <- as.factor(n_data$Education)
```

##constructing the fake variables for the attribute "education."

```{r}
fake<- dummyVars(~.,data=n_data)
univ_data <- as.data.frame(predict(fake,n_data))
```

##Since we must run the function once more, we must first generate the seed and split the data into groups for training (60%) and validation (40%).

```{r}
set.seed(1)
t_data <- sample(row.names(univ_data), 0.6*dim(univ_data)[1])
valids_data <- setdiff(row.names(univ_data),t_data)
train <- univ_data[t_data,]
valid <- univ_data[valids_data,]
t(t(names(train)))
```
```{r}
summary(train)
```
```{r}
cat("The size of the training dataset is:",nrow(train))
```
```{r}
summary(valid)
```
```{r}
cat("The size of the validation dataset is:",nrow(valid))
```
Now let's standardize the information.

```{r}
t_norm <- train[,-10]
v_norm <- valid[,-10]
norm <- preProcess(train[,-10],method=c("center","scale"))
5
```

```{r}
t_norm <- predict(norm,train[,-10])
v_norm <- predict(norm,valid[,-10])
```

## Question:

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 =
1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and
Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code
using k = 1. Remember to transform categorical predictors with more than two categories
into dummy variables first. Specify the success class as 1 (loan acceptance), and use the
default cutoff value of 0.5. How would this customer be classified?


```{r}
latest.customer <- data.frame(
Age = 40,
Experience = 10,
Income = 84,
Family = 2,
CCAvg = 2,
Education.1 = 0,
Education.2 = 1,
Education.3 = 0,
Mortgage = 0,
Securities.Account = 0,
CD.Account = 0,
Online = 1,
CreditCard = 1
)

# Normalize the new customer dataset
cust_latest.norm <- predict(norm, latest.customer)
```

Applying kNN classification
```{r}
prediction <- class::knn(train = t_norm,
test = cust_latest.norm,
cl = train$Personal.Loan, k = 1)
prediction
```
2. What is a choice of k that balances between overfitting and ignoring the predictor information?

```{r}
# Determine each value of k's accuracy.
#Determine the k-value range that will be taken into account.
accuracy <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
kn <- class::knn(train = t_norm,
test = v_norm,
cl = train$Personal.Loan, k = i)
accuracy[i, 2] <- confusionMatrix(kn,
as.factor(valid$Personal.Loan),positive = "1")$overall[1]
}
which(accuracy[,2] == max(accuracy[,2]))
```
```{r}
accuracy
```
Three is the best-performing k value between 1 and 15. Because it strikes a compromise between overfitting and disregarding forecasts, this k is the most accurate for 3.

```{r}
plot(accuracy$k,accuracy$overallaccuracy)
```

3. Show the confusion matrix for the validation data that results from using the best k.

```{r}
predicts<- class::knn(train = t_norm,
test = v_norm,
cl = train$Personal.Loan, k=3)
confusionMatrix(predicts,as.factor(valid$Personal.Loan))
```
4. Consider the following customer: Age = 40, Experience = 10, Income = 84,
Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0,
Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit
Card = 1. Classify the customer using the best k.


```{r}
latest.cust2 <- data.frame(
Age = 40,
Experience = 10,
Income = 84,
Family = 2,
CCAvg = 2,
Education.1 = 0,
Education.2 = 1,
Education.3 = 0,
Mortgage = 0,
Securities.Account = 0,
CD.Account = 0,
Online = 1,
CreditCard = 1)

#Normalizing the 2nd client dataset
latest.cust2_norm <- predict(norm,latest.cust2)
```

5. Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply
the k-NN method with the k chosen above. Compare the confusion matrix of the test set
with that of the training and validation sets. Comment on the differences and their reason.

```{r}
set.seed(500)
T_Index <- sample(row.names(univ_data), .5*dim(univ_data)[1])#create train index
9
```
```{r}
#create validation index
V_Index <- sample(setdiff(row.names(univ_data),T_Index),.3*dim(univ_data)[1])
Test_Index =setdiff(row.names(univ_data),union(T_Index,V_Index))#create test index
train.df <- univ_data[T_Index,]
cat("The size of the new training dataset is:", nrow(train.df))
```
```{r}
valid.df <- univ_data[V_Index, ]
cat("The size of the new validation dataset is:", nrow(valid.df))
```
```{r}
test.df <- univ_data[Test_Index, ]
cat("The size of the new test dataset is:", nrow(test.df))
```
```{r}
n_values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.df.norm <- predict(norm, train.df[, -10])
valid.df.norm <- predict(norm, valid.df[, -10])
test.df.norm <- predict(norm ,test.df[,-10])
```

Using training, testing, and validation data, kNN processing and confusion matrix generation are performed.

```{r}
predict_3 <- class::knn(train = train.df.norm,
test = test.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(predict_3,as.factor(test.df$Personal.Loan))
```
```{r}
predict_4 <- class::knn(train = train.df.norm,
test = valid.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(predict_4,as.factor(valid.df$Personal.Loan))
```
```{r}
predict_5 <- class::knn(train = train.df.norm,
test = train.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(predict_5,as.factor(train.df$Personal.Loan))
```

