---
title: "FML Assignment 3"
output: html_document
date: "2024-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
```{r}
library(reshape)
```
```{r}
library(reshape2)
```

```{r}
library(caret)
```
```{r}
library(e1071)

## Loading the data csv file
U_B.df <- read.csv("C:/Users/TEJA/Downloads/UniversalBank (1).csv")

head(U_B.df)
```
```{r}
tail(U_B.df)
```
```{r}
## defining Column names
colnames(U_B.df)
```
```{r}
#converting information into factors using the categories.

U_B.df$`Personal Loan` = as.factor(U_B.df$Personal.Loan)
U_B.df$Online = as.factor(U_B.df$Online)
U_B.df$CreditCard = as.factor(U_B.df$CreditCard)
```

```{r}
#dividing the data in to two halves, with 40% going into the validation set and 60% going into the training set
set.seed(456)

# 60 % training data and 40% Validation data respectively
U_B.df.train.data <- sample(row.names(U_B.df), 0.6*dim(U_B.df)[1])
U_B.df.valid.data <- setdiff(row.names(U_B.df), U_B.df.train.data)

# assigning training data and validation data to a dataframe from dataset respectively
U_B.df.train <- U_B.df[U_B.df.train.data, ]
U_B.df.valid <- U_B.df[U_B.df.valid.data, ]

# Duplicating the data frame UB.train
train <- U_B.df[U_B.df.train.data, ]
# Duplicating the data frame UB.valid
valid = U_B.df[U_B.df.train.data,]
```


##A. Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table().

```{r}

## melt data from data
melted.df = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")

```

```{r}

# Casting the melted data
cast.df <- dcast(melted.df, CreditCard + Personal.Loan ~ value, fun.aggregate = length)

# Casting column number 14: Personal loan, ID, and credit card, respectively
cast.df[,c(1,2,3,14)]

```

#B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].

```{r}

#According to the pivot table, the value for Loan is 89, and the value for CC is 

UB.df.Loan.CC1 <- 89/3000

# which is 29 %.
UB.df.Loan.CC1

```
##C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.


```{r}

 ## transforming the train data frame into a lengthy format, using "Online" as a variable to be melted and "Personal.Loan" as an identification
melt1.df = melt(train,id=c("Personal.Loan"),variable = "Online")

```

```{r}

## creating a long format out of the train data frame, using "CreditCard" as an identifier and "Online" as a variable that needs to be melted
melt2.df = melt(train,id=c("CreditCard"),variable = "Online")

```

```{r}

# Casting Personal loan and online values
# Assuming both Personal.Loan and Online are binary (0 or 1)
cast1.df =dcast(melt1.df,`Personal.Loan`~Online)

```

```{r}

# Casting Personal loan and online values
cast2.df=dcast(melt2.df,CreditCard~Online)

```

```{r}

#shows the number of personal loans in reference to online
UB.df.Loanonline=cast1.df[,c(1,13)]
UB.df.LoanCC = cast2.df[,c(1,14)]
UB.df.Loanonline

```

```{r}

# shows the number of credit cards in reference to Online.
UB.df.LoanCC

```

##D. Compute the following quantities [P (A | B) means “the probability of A given B”]:#1.P (CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors) #2.P(Online=1|Loan=1) #3.P (Loan = 1) (the proportion of loan acceptors) #4.P(CC=1|Loan=0) #5.P(Online=1|Loan=0) #6.P(Loan=0)

```{r}

# creating a pivot table with the columns 14 and 10 representing personal loan
table(train[,c(14,10)])

```

```{r}

# Creating a pivot table for column 13 and 10 which is online and personal loan

table(train[,c(13,10)])

```

```{r}

# Personal loan pivot table There are 2725 and 275 from training, respectively
table(train[,c(10)])

```


1.P (CC = 1 | Loan = 1

```{r}

## We can obtain the CC= 1 and Loan = 1 values by referring to the above p
UB.df.CCUB.Loan1 = 89/(89+200)
UB.df.CCUB.Loan1

```

2.P(Online=1|Loan=1)

```{r}

##We can get the online = 1 and loan = 1 values from the pivot table above UB.ONUB.Loan1
UB.df.ONUB.df.Loan1 =177/(177+112)

```


3.P(Loan = 1)

```{r}

 # By referring the above pivot table we can get the Loan = 1
UB.df.Loan1 =289/(289+2711)
UB.df.Loan1

```

4.P(CC=1|Loan=0)

```{r}

#Using the pivot table above, we can obtain the CC = 1 and Loan = 0 values
UB.df.CCLoan.01= 794/(794+1917) 
UB.df.CCLoan.01

```

5.P(Online=1|Loan=0)

```{r}

# We can get the online = 1 and loan = 0 values from the pivot table above.
UB.df.ON1.L0= 1665/(1665+1046)
UB.df.ON1.L0

```

6.P(Loan=0)

```{r}

# We can obtain the Loan = 0 values by the pivot table above.
UB.df.Loan0= 2711/(2711+289)
UB.df.Loan0

```

##E. Use the quantities computed above to compute the naive Ba1 probability P(Loan = 1 | CC = 1, Online = 1).

```{r}

# Given probabilities
Probability_CC_given_Loan1 <- 0.096
Probability_Online_given_Loan1 <- 0.833
Probabilities_Loan1 <- 0.0125

```

```{r}

# Calculate Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1)
#Universal_Bank_Naivebayes <- (P_CC_given_Loan1 * P_Online_given_Loan1 * P_Loan1) / (P_CC * P_Online)

UB.df_Naivebayes <- (Probability_CC_given_Loan1)*(Probability_Online_given_Loan1)*(Probabilities_Loan1)
UB.df_Naivebayes

```

##F. Compare this value with the one obtained from the pivot table in (b). Which is a more accurate estimate?

```{r}

# Naive Bayes Probability (from calculation in E)
naive_bayes_probability <- 0.0009996

# Pivot Table Probability (from part B)
pivot_table_probability <- 0.02966667

# Comparing the probabilities and print a message indicating which is more accurate
if (naive_bayes_probability > pivot_table_probability) {
  message("Naive Bayes Probability is more accurate: ", naive_bayes_probability)
} else if (naive_bayes_probability < pivot_table_probability) {
  message("Pivot Table Probability is more accurate: ", pivot_table_probability)
} else {
  message("Both probabilities are the same: ", naive_bayes_probability)
}

```

Based on the comparison, the pivot table probability (0.02966667) is considered more accurate compared to the Naive Bayes probability.

##G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}

names(U_B.df)

```

```{r}

names(U_B.df.train)

```

```{r}

# Selecting the relevant columns for training
U_B.df.train <- U_B.df.train[, c("CreditCard", "Online", "Personal Loan")]

# Renamimg the columns to remove spaces
colnames(U_B.df.train) <- c("CreditCard", "Online", "PersonalLoan")

# Converting 'CreditCard' and 'Online' to factors with appropriate levels
U_B.df.train$CreditCard <- factor(U_B.df.train$CreditCard, levels = c(0, 1), labels = c("No", "Yes"))
U_B.df.train$Online <- factor(U_B.df.train$Online, levels = c(0, 1), labels = c("No", "Yes"))

```


##Print the probability

```{r}

print("Probability of Loan = 1 given CC = 1 and Online = 1:")

```














